Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-21 manuscript/manuscript.Rmd
partially got the first figure to work as r code

@@

```{r}
library(tidyverse)
msy <- read_csv("../data/msy.csv.gz")
ppo <- read_csv("../data/ppo-200.csv")
escapement <- read_csv("../data/escapement.csv.gz", show_col_types = FALSE)

# rename some columns for figure purposes:
colnames(msy) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z')
colnames(ppo) <- c('1', 't', 'X', 'Y', 'Z', 'action', 'reward', 'rep')

# ppo dataset uses step rewards rather than cumulative rewards, which is what is useful here.
# so, switch stuff around:
ppo |> 
    group_by(rep) |> 
    mutate(cumulative_reward = cumsum(reward)) |>
    within(rm("reward", "1")) |>
    rename(reward = cumulative_reward)

@@

i <- which.min(abs(actions - best$action* 0.8))
tac_sim <- msy |> 
    filter(action == actions[[i]]) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "tac")

ppo_sim <- ppo |> 
    mutate(escapement = (X+1) - action * (X+1),
        effort = action) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(policy = "ppo") # already in natural units

best_e <- 0.62 # optimal constant escapement, computed from ../data/escapement.csv.gz
esc_sim <- escapement |> 
    rename(escapement = action) |>
    filter(escapement == best_e) |> # data had a naming problem
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1)
```

Our results suggest that DRL can find solutions that are more sustainable *and* more profitable than both MSY and TAC.
Indeed, DRL finds solutions in which, for certain periods of time, no harvest is performed on the system.
These periods effectively allow the system to, in a sense, "reset" itself.
They lead to a distinctive oscillating behavior of the system state over time, with periods of relative depletion followed by periods of growth. 
Figure \ref{fig:tbd} exemplifies this point by comparing one times series of the system state obtained using each of these policies.

```{r fig.cap="TBD."}

```

>>>>

```{r}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, 
                      warning = FALSE, fig.height=5, fig.width=7.5)
library(tidyverse)
```

@@

```{r}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, 
                      warning = FALSE, fig.height=5, fig.width=7.5)
msy <- read_csv("../data/msy.csv.gz")
ppo <- read_csv("../data/PPO250.csv.gz")
escapement <- read_csv("../data/escapement.csv.gz", show_col_types = FALSE)

# rename some columns for figure purposes:
colnames(msy) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z')
colnames(ppo) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z') 
colnames(escapement) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z')

# ppo dataset uses step rewards rather than cumulative rewards, which is what is useful here.
# so, switch stuff around:
#ppo |> 
#    group_by(rep) |> 
#    mutate(cumulative_reward = cumsum(reward)) |>
#    within(rm("reward", "1")) |>
#    rename(reward = cumulative_reward)

@@

actions <- unique(msy$action)
i <- which.min(abs(actions - best_action* 0.8))
tac_sim <- msy |> 
    filter(action == actions[[i]]) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "tac")

ppo_sim <- ppo |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "ppo")

best_e <- 0.62 # optimal constant escapement, computed from ../data/escapement.csv.gz
esc_sim <- escapement |> 
    rename(escapement = action) |> # data had a naming problem
    filter(escapement == best_e) |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1)

msy_tac_ppo <- bind_rows(msy_sim, tac_sim, ppo_sim)
```

Our results suggest that DRL can find solutions that are more sustainable *and* more profitable than any constant mortality strategy.
Indeed, DRL finds solutions in which, for certain periods of time, no harvest is performed on the system.
These periods effectively allow the system to, in a sense, "reset" itself.
They lead to a distinctive oscillating behavior of the system state over time, with periods of relative depletion followed by periods of growth. 
Figure \ref{fig:tbd} exemplifies this point by comparing one times series of the system state obtained using each of these policies.

```{r fig.cap="TBD."}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, 
                      warning = FALSE, fig.height=5, fig.width=7.5)
policy_labels <- list("msy" = "Opt. Const. Mortality", "tac"= "80% Opt. Const. Mortality", "ppo"= "DRL")
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 0) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) + 
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/5)
```
